---
title: "April 7 Lab"
author: "Jordy Taylor"
date: "Tuesday, April 07, 2015"
output: html_document
---

ANCOVA: 
------------------------------
DOWNLOAD ACORN DATASET, VARIABLES WITH SPECIES, REGION, AREA OR RANGE, ACORN SIZE AND TREE HEIGHT

TEST WHETHER THE SLOPE THAT RELATES ACORN SIZE TO RANGE DIFFERS FOR THE DIFFERENT REGIONS (atlantic versus california). If there is no difference in slope, is there a difference in range size in the 2 regions when you take acorn size into account?

ACORN SIZE TO RANGE: For Atlantic and Califoria 
```{r}
acorns <- read.delim("C:/Users/Jordy/Desktop/Biometry Lecture Notes/acorns.dat")
View(acorns)

Atlantic<-acorns[acorns$Region=="Atlantic",]
Atlantic
California<-acorns[acorns$Region=="California",]
California


#For Atlantic  and Cali

Atl<-lm(Acorn_size~Range, data=acorns, subset=Region=="Atlantic")
Cali<-lm(Acorn_size~Range, data=acorns, subset=Region=="California")

plot(Acorn_size~Range, data=acorns)  #VERY DIFFERENT SLOPES

abline(coef(Atl), lty=1)
abline(coef(Cali), lty=2)

#AOV
aov.acorn<-(aov(Acorn_size~Range*Region, data=acorns))  #significant effect of region 
aov.acorn

residualsacorns<-resid(aov.acorn)
residualsacorns
fit.acorn<-fitted(aov.acorn)
fit.acorn

plot(aov.acorn)

#Now log transform it
log.aov.acorn<-(aov(log(Acorn_size)~log(Range)*Region, data=acorns))
log.aov.acorn

plot(log.aov.acorn)  #variance is not equal

#Now i just want to run this just as a function of just the size logged

log.size<-aov(log(Acorn_size)~Range*Region, data=acorns) #this is with the interaction
plot(log.size)
summary(log.size)

log.size.noi<-aov(log(Acorn_size)~Range+Region, data=acorns)  #this is with no interaction
plot(log.size.noi)
summary(log.size.noi)

#so range does not really have an effect on the acorn size, rather, there is an effect of region on the acorn size (as shown by the P value)


```

Now look to see if there is an effect with tree height instead of acorn size

```{r}

Atlantic<-lm(Tree_Height~Range, data=acorns, subset=Region=="Atlantic")
California<-lm(Tree_Height~Range, data=acorns, subset=Region=="California")

plot(Tree_Height~Range, data=acorns)  #VERY DIFFERENT SLOPES

abline(coef(Atlantic), lty=1)
abline(coef(California), lty=2)

#AOV
aov.acorn.height<-(aov(Tree_Height~Range*Region, data=acorns))  #significant effect of region 
aov.acorn.height

residualstree<-resid(aov.acorn.height)
residualstree
fit.tree<-fitted(aov.acorn.height)
fit.tree

plot(aov.acorn.height)

#Now log transform it
log.aov.tree<-(aov(log(Tree_Height)~log(Range)*Region, data=acorns))
log.aov.tree

plot(log.aov.tree)  #variance is not equal

#Now i just want to run this just as a function of just the size logged

tree.interaction<-aov(Tree_Height~Range*Region, data=acorns) #this is with the interaction

plot(tree.interaction)
summary(tree.interaction)

tree.no.interaction<-aov(Tree_Height~Range+Region, data=acorns)  #this is with no interaction
plot(tree.no.interaction)
summary(tree.no.interaction)


```

UNEQUAL VARIANCE FOR SOME SQUID

```{r}
squid <- read.csv("C:/Users/Jordy/Desktop/Biometry Lecture Notes/squid.csv")
View(squid)

library(lattice)
xyplot(Testisweight~DML|MONTH, data=squid)


boxplot(Testisweight~MONTH, data=squid)

#Do a normal Ancova to see how data fit assumption with the interaction

interaction<-aov(Testisweight~DML*as.factor(MONTH), data=squid)
plot(interaction)
summary(interaction)


#DO A NORMAL ANCOVA TO SEE HOW THE DATA FIT WITHOUT THE INTERATION
no.interaction<-aov(Testisweight~DML+as.factor(MONTH), data=squid)
plot(no.interaction)
summary(no.interaction)
```

Log Transform Both X and y

```{r}
#well it still doesn't look nice, so log transform the data to see if we can get a better fitting model

log.interaction<-aov(log(Testisweight)~log(DML)*as.factor(MONTH), data=squid)
plot(log.interaction)
summary(log.interaction)


#DO A NORMAL ANCOVA TO SEE HOW THE DATA FIT WITHOUT THE INTERATION
log.no.interaction<-aov(log(Testisweight)~log(DML)+as.factor(MONTH), data=squid)
plot(log.no.interaction)
summary(log.no.interaction)


```

All of this data looks like shit, so you want to normalize the residuals

```{r}

library(nlme)
fit.reg<-gls(Testisweight~DML, data=squid)
summary(fit.reg)

var.fix<-varFixed(~DML)
fit.var<-gls(Testisweight~DML, weights=var.fix, data=squid)
summary(fit.var)

anova(fit.reg, fit.var)
#smaller AIC is better so the one with the VARIATION is better fitting the  model

```

So now try a differnt model to see if it will fit better: exponential

```{r}
var.fix
fit.var<-gls(Testisweight~DML, weights=var.fix, data=squid)

varianexp<-varExp(form=~DML)
fit.exp<-gls(Testisweight~DML, weights=varianexp, data=squid)
summary(fit.exp)
anova(fit.var, fit.exp)

#the exponential is lower so it fits better

```

NOw try the power model
```{r}
varianexp<-varExp(form=~DML)
fit.exp<-gls(Testisweight~DML, weights=varianexp, data=squid)

vpow<-varPower(form=~DML)
fit.pow<-gls(Testisweight~DML, weights=vpow, data=squid)

anova(fit.exp, fit.pow)

#So the power one fits even better than all the other models that we just looked at

#Here i want to identify how there is difference variance within my different groups

varid<-varIdent(form=~1|MONTH)
fit.id<-gls(sqrt(Testisweight)~as.factor(MONTH), data=squid)

summary(fit.id)
anova(fit.id)   #month is significantly importnat here because null is not 0
```


```{r}
varC<-varComb(varPower(form=~DML), varExp(form=~DML), varFixed(~DML))
fit.com<-gls(Testisweight~DML, data=squid, weights=varC)
summmary(fit.com)
anova(fit.com, fit.pow)


#These mutha lickers are the exact same model. WTF.