March 10th Lab
Two-sample test

Take the data in 'cloud.csv

```{r}
clouds <- read.csv("C:/Users/Jordy/Desktop/Biometry Lecture Notes/clouds.csv")
View(clouds)
aov(Unseeded~Seeded, data=clouds)
plot(aov(Unseeded~Seeded, data=clouds))

qqnorm(clouds)


```
Determine if data are distributed normally. If not find a good transformation and use a
parametric test

```{r}
plot(hist(clouds$Unseeded))
plot(hist(clouds$Seeded))

hist(log(clouds$Unseeded))
hist(log(clouds$Seeded))

plot(clouds)
plot(log(clouds))

aov(log(Unseeded)~log(Seeded),data=clouds)

summary(aov(log(Unseeded)~log(Seeded),data=clouds))

plot(aov(log(Unseeded)~log(Seeded),data=clouds))

```




perform a Wilcoxon test on the untransformed data
```{r}
x<-(clouds$Unseeded)
x
y<-(clouds$Seeded)
    
xy<- c(x,y)    

ranks<-rank(xy)
xranks<-ranks[1:length(x)]
yranks<-ranks[(length(x)+1) : (length(x) + length(y))]
sum.R<-sum(yranks)
    
xranks    
    
yranks    

c<-length(x)*length(y)+ (length(y)*length(y)+1)/2-sum.R
c   #190.5

if ((length(x)*length(y)-c)<c)
  f
u<-c
else
  f
u<-length (x)*length(y)-c
u   #485.5
(1-pwilcox(u,length(x),length(y)))

#  u>c

Other way to do it:
 
wilcox.test(x,y)

```

One-way anova analogue


```{r}
cuckoo <- read.csv("C:/Users/Jordy/Desktop/Biometry Lecture Notes/cuckoo.csv")
>   View(cuckoo)
m.p<-(cuckoo[cuckoo$species=="m.pipit",])
t.p<-(cuckoo[cuckoo$species=="t.pipit",])
sprw<-(cuckoo[cuckoo$species=="h.sprw",])
robin<-(cuckoo[cuckoo$species=="robin",])
wtail<-(cuckoo[cuckoo$species=="wtail",])
wren<-(cuckoo[cuckoo$species=="wren",])

hist(m.p$length)
hist(t.p$length)
hist(sprw$length)
hist(robin$length)
hist(wtail$length)
hist(wren$length)

as.factor(cuckoo$species)
summary(aov(log(length)~species, data=cuckoo))

plot(aov(log(length)~species, data=cuckoo))
#this has more power than the kruskal test, this shows that the data are significantly related to one another


library(lattice)
histogram(~length|species, data=cuckoo)

kruskal.test(length~species, data=cuckoo)

#this says that these data are not significanlty correlated
     
```

 Planned comparison
 
  Test the a priori hypothesis that the control
is different than the 4 treatments. Make sure you test to see if there is any effect first.

```{r}
pea <- read.csv("C:/Users/Jordy/Desktop/Biometry Lecture Notes/pea.csv")
>   View(pea)

control<-(pea[pea$sugar=="cntrl",2])
glu<-(pea[pea$sugar=="glu",2])
fru<-((pea[pea$sugar=="fru",2]))
gf<-(pea[pea$sugar=="glufru",2])
suc<-(pea[pea$sugar=="suc",2])

L<-mean(control)-.25*mean(glu)-.25*mean(fru)-.25*mean(gf)-.25*mean(suc)


L  #this is the difference of the control mean from the means of the experiemental groups. See function a posteriori function below too look at comparisons

SEl<-sqrt(1+.25^2+.25^2+.25^2+.25^2)* sqrt (5.4289/10)
SEl
#L-0/SEl
10.2/.82
#t=12.43902
1-pt(12.43902,df=45)
#p=2.22 x 10^-16

with(pea,pairwise.t.test(length,sugar))


as.factor(pea$sugar)
aov(length~sugar, data=pea)
summary(aov(length~sugar, data=pea))
plot(aov(length~sugar, data=pea))#normal




```

Unplanned comparisons

Use pairwise t-tests and Tukey's HSD test to perform a post hoc test on the Cuckoo data present in
'cuckoo.csv'

```{r}
View(cuckoo)

with(cuckoo,pairwise.t.test(length,species))


as.factor(cuckoo$species)
summary(aov(log(length)~species, data=cuckoo))

TukeyHSD(aov(length~as.factor(species), data=cuckoo))
par(las=1)
par(mar=c(5,13,4,2))
plot(TukeyHSD(aov(length~as.factor(species), data=cuckoo)))  

```
```














