---
title: 'march31: Regression and Kendall'
author: "Jordy Taylor"
date: "Tuesday, March 31, 2015"
output: html_document
---
1: Functions
```{r}
add3<-function(a,b,c)
  {
  a+b+c
  }
#now call the function with some data
add3(1, 10, .05)
#11.05
```

Write a function call t.tail that always returns the area under the curve in the two tails of a t distribution. So the integral of the values of the tails of a distribution.

```{r}
x<-seq(8,28)
plot(dnorm(x,mean=18, sd=3.162278)~x)
t.tail<-function(arg)
  

```


2. Non-Parametric

```{r}
acorns <- read.delim("C:/Users/Jordy/Desktop/Biometry Lecture Notes/acorns.dat")
View(acorns)

Atlantic<-acorns[acorns$Region=="Atlantic",]
Atlantic
California<-acorns[acorns$Region=="California",]
California


lm(Acorn_size~Range, data=Atlantic)
plot(Acorn_size~Range, data=Atlantic)
abline(coef(lm(Acorn_size~Range, data=Atlantic)))
summary(lm(Acorn_size~Range, data=Atlantic))
#.1499 of the variation of Range to Acorn size in the atlantic


#Residuals for Atlantic:
A<-lm(Acorn_size~Range, data=Atlantic)
plot(resid(A)~fitted(A))
plot(A)

#California data

lm(Acorn_size~Range, data=California)
plot(Acorn_size~Range, data=California)
abline(coef(lm(Acorn_size~Range, data=California)))
summary(lm(Acorn_size~Range, data=California))


#Residuals for California:

C<-lm(Acorn_size~Range, data=California)
plot(resid(C)~California$Acorn_size)



```
Log Transform this data:

```{r}  
#Atlantic
lm.h<-lm(log(Acorn_size)~Range, data=Atlantic)
plot(log(Acorn_size)~Range, data=Atlantic)
abline(coef(lm.h))
summary(lm.h)
plot(lm.h)
#.2942

#Residuals:
plot(resid(lm.h)~Atlantic$Range)
plot(resid(lm.h)~fitted(lm.h))

#California
lm.c<-lm(log(Acorn_size)~Range, data=California)
plot(log(Acorn_size)~Range, data=California)
abline(coef(lm.c))
summary(lm.c)
#.01495

#Residuals:
plot(resid(lm.c)~California$Range)

```



Use Kendall's Robut Regressions Technique to estimate the slope and intercept:

```{r}
x<-Atlantic$Range
y<-Atlantic$Acorn_size

kendall.robust<-function(x,y)
  {
  Svec<-NULL
  for (i in 1: length(x))
    {
    for (j in 1:length(y))
      {
      if (j>i)
        {
        Svec<-c(Svec,(y[j]-y[i])/(x[j]-x[i]))
        }
      }
    }
  b1<-median(Svec)
  b1
bo<-median(y-b1*x)
bo
c(bo,b1)
  }

kendall.robust(Atlantic$Range,Atlantic$Acorn_size)

cor.test(y, x, method="kendall")

#data:  y and x
#z = 3.4155, p-value = 0.0006367
#alternative hypothesis: true tau is not equal to 0
#sample estimates:
#      tau 
#0.4662342

```





3. Multiple Regression: test whether calcium concentration affects the amount of mercury in muscle.

```{r}

bass <- read.delim("C:/Users/Jordy/Desktop/Biometry Lecture Notes/bass.dat")
View(bass)

lm.calcium<-lm(log(Avg_Mercury)~Calcium, data=bass)
lm.calcium
plot(log(Avg_Mercury)~Calcium, data=bass)
abline(coef(lm.calcium))

plot(resid(lm.calcium)~bass$Calcium)

summary(lm.calcium)
#Variability explain= 27%

```

Now add in Alkalinity, Chlorophyll, and pH as additional terms.

```{r}
#Effect of Calcium and Alkalinity:

lm.calc.alk<-lm(log(Avg_Mercury)~Calcium+Alkalinity, data=bass)
lm.calc.alk
plot(log(Avg_Mercury)~Calcium+Alkalinity, data=bass)
abline(coef(lm.calc.alk))
pairs(bass[,3:7])

summary(lm.calc.alk)
#.5045



#Effect of Calcium, Alkalinity, and Chlorophyll

lm.calc.alk.chlo<-lm(log(Avg_Mercury)~Calcium+Alkalinity+Chlorophyll, data=bass)
lm.calc.alk.chlo
plot(log(Avg_Mercury)~Calcium+Alkalinity+Chlorophyll, data=bass)
summary(lm.calc.alk.chlo)

#Effect of Everything on Mercury

lm.calc.alk.chlo.ph<-lm(log(Avg_Mercury)~Calcium+Alkalinity+Chlorophyll+pH, data=bass)

lm.calc.alk.chlo.ph
plot(log(Avg_Mercury)~Calcium+Alkalinity+Chlorophyll+pH, data=bass)
summary(lm.calc.alk.chlo.ph)




```

When you look at the different plots yielded by each added variable, you see that the independent variable has some sort of a different effect on the mercury content found. It looks as if the Alkalinity/pH have a significant effect on the Mercury content along with the chlorophyll activity. You can see a linear relationship between the pH/alkalinity on the effect of mercury and with the chlorophyll activity, you can see a clumped distribution of the dependent variable. More than one variable can effect the dependent variable, where 2 independent variables will have a greater effect than just 1 independent variable. 




Phenotypes Data

```{r}
library(dplyr)
library("dplyr", lib.loc="~/R/win-library/3.1")
setwd("C:/Users/Jordy/Desktop/Biometry Lecture Notes")
phen<- read.csv("phenotypes2015-03-30.csv")
phen
names(phen)
dim(phen)
tmp<-phen%


phenotypes2015 <- read.csv("C:/Users/Jordy/Desktop/Biometry Lecture Notes/phenotypes2015.csv")
View(phenotypes2015)

#filter allows you to select just rows so 1a1 and 1a2, so we are reducing to just the rows 1a1 and 1a2. 


tmp<-phen%>%select(Accession, treatment)
tmp
tmp[1:10,]
tmp<-phen[,c("Accession","treatment")]
tmp

#summarize does the summary statistics, so this gives ou the average fruit number and avg. fruit length for all the accession numbers for 1a1 and 1a2.


tmp2<-phen%>%filter(experiment%in%c("1a1","1a2"))%>%group_by(Accession)
tmp2


tmp3<-tmp2%>%summarise(meanfruit=mean(fruitnum,na.rm=T), meanlength=mean(avg.fruit.length, na.rm=T))

tmp3

plot(meanfruit~meanlength, data=tmp3)
regression<-lm(meanfruit~meanlength, data=tmp3)
abline(coef(regression))
summary(regression)

#coef gives you the y intercept and slope to calculate the abline



```




1 Regression using likelihood
See if you can use the code in the regression likelihood lecture to estimate the relationship between
log(body size) and log(brain size) from the 'Animals' data set found in the MASS library. Use
log(body size) as the independent (X) variable.
Check your estimates using 'lm()'.
For fun, try to use the function probf.plot instead of probf in the optim statement. This will
animate the curve-fitting process.

```{r}
library("MASS", lib.loc="C:/Program Files/R/R-3.1.2/library")
View(Animals)

body<-c(Animals$body)
body
brain<-c(Animals$brain)
brain





lm.bodyonbrain<-lm(logbody~logbrain)
lm.bodyonbrain

plot(log(body)~log(brain), data=Animals)
abline(coef(lm.bodyonbrain))



probf <- function(p,X,Y)
{
  b0 <- p[1]; b1 <- p[2]; sd <- p[3]

  rsd <- (b0+b1*X)-Y
  -sum(dnorm(rsd,mean=0,sd=sd,log=T))
}

X=body
Y=brain

optim(c(1,0,5),probf,X=log(body),Y=log(brain))




probf.plot <- function(p,X,Y)
{
  b0 <- p[1]; b1 <- p[2]; sd <- p[3]
  rsd <- (b0+b1*X)-Y
  def.par <- par(no.readonly = TRUE)
  layout(matrix(c(1,1,2,2),2,2,byrow=T),respect=T)
  plot(Y~X,main=paste("Ln Likelihood:",sum(dnorm(rsd,mean=0,sd=sd,log=T)),"\nY = ",round(b0,2),"+",round(b1,4),"X"))
  abline(c(b0,b1))
  tmp=apply(cbind(X,Y,rsd),1,function(x,rsd){
    points(c(x[1],x[1]),c(x[2],(x[2]+x[3])),type="l",
           col=ifelse(x[3]<0,"green","red"))
  },rsd=rsd)
  hist(rsd,freq=F,main=paste("Distribution of rsd\nmu=0, sigma=",sd))
  rug(rsd,col="red")
  pts <- seq(min(rsd)*2,max(rsd)*2,length=100)
  points(dnorm(pts,0,sd)~pts,type="l",col="red")
  par(def.par)
  Sys.sleep(0.5)
  -sum(dnorm(rsd,mean=0,sd=sd,log=T))
} 




probf.plot(c(1,0,5), X=log(body), Y=log(brain))

optim(c(1,0),probf.plot, X=log(body), Y=log(brain), data=Animals)


```






