---
title: "Regression Lab"
author: "Jordy Taylor"
date: "Tuesday, March 24, 2015"
output: html_document
---


Take the data in fruit_count.csv.  Use the data in this dataframe to examine the hypothesized relationship. Use the linear regression techniques mentioned in class. Calculate the necessary quantities in R but don't use the lm . Make sure to plot your data and the resulting regression line (function abline() takes the intercept and slope in a veotr: 'abline(c(b0,b1))

```{r}
fruit_count <- read.csv("C:/Users/Jordy/Desktop/Biometry Lecture Notes/fruit_count.csv")

fruit_count=fruit_count[order(fruit_count$diameter),]
#View(fruit_count)
#diameter<-c(65,78.5,76,47,25,28,10,22,40,54,73,44)
diameter=fruit_count$diameter
#fruits<-c(2027,1879,2723,984,442,519,347,666,1882,1111,1669,1193)
fruits=fruit_count$fruits
plot(fruits~diameter)


#SLOPE
xdev<-diameter-mean(diameter)
ydev<-fruits-mean(fruits)
slope<-sum((xdev)*(ydev))/sum(xdev^2)
slope
#28.01765

#Yintercept
y<-mean(fruits)-mean(diameter)*slope
y
#-26.49

plot(fruits~diameter)
abline(c(y,slope))



```
Now calculate the confidence interval around b1 for the fruit count

```{r}
y.hat <- y + slope * diameter
ss.explained <- sum((y.hat-mean(fruits))^2)
ss.unexplained <- sum((y.hat-fruits)^2)
stderr.b1 <- sqrt((ss.unexplained/7)/sum(xdev^2))
stderr.b1
#6.344425

lower <- slope - qt(0.95,10)*stderr.b1
upper <- slope + qt(0.95,10)*stderr.b1
lower
#16.51863
upper
#39.51667
```
Confidence Intervals around Y
```{r}
MS.exp<-ss.explained/1
MS.error<-ss.unexplained/10
MS.exp/MS.error
#27.85998

1-pf((MS.exp/MS.error),1,10)
#0.0003585396

std.yhat <- sqrt((ss.unexplained/10)*((1/12)+(xdev^2/sum(xdev^2))))
std.yhat


upper <- y.hat + qt(0.95,10)*std.yhat
upper  
lower <- y.hat - qt(0.95,10)*std.yhat
lower  

points(upper~diameter,type="l",lty=2)
points(lower~diameter,type="l",lty=2)

```
 R Squared
 
```{r}
R.squared <- ss.explained/(ss.explained+ss.unexplained)
R.squared
#.73586
lm.summary<- summary(lm(fruits~diameter))
lm.summary
lm.summary$r.squared

#.7358689

